---
title: "Stat363 - HW3"
author: "Mert Göksel"
date: "1/3/2022"
output: pdf_document
---

#Q2:

##A:

```{r, message=FALSE}
library(openxlsx)
setwd("~/Stat/R/Stat363/")
grocery <- read.xlsx("./groceryretailer.xlsx")
pairs(grocery)
```

x3 is a categorical variable and no linear relationship can be seen in any 
scatter plot

```{r, message=FALSE}
library(corrplot)
corrplot(cor(grocery))
```

y and x3 is highly correlated, x1 and x2 correlation to y is lower. 
No multicolinearity is detected, the only high correlation is between x3 and y
but x3 is a categorical value while y is a numerical one.

##B:

```{r, message=FALSE}
model <- lm(y ~ ., data = grocery)
library(car)
vif(model)
#1.008596 1.019598 1.014364 for reference.

library(dplyr)
cor(grocery)
```

We apply unit length scaling to grocery data and get the correlation matrix

first of all get the sum of squares
$$\sum (x_i - \bar x)^2$$ 
for each column

```{r}
s11 <- sum((grocery$x1 - mean(grocery$x1))^2) 
s22 <- sum((grocery$x2 - mean(grocery$x2))^2) 
s33 <- sum((grocery$x3 - mean(grocery$x3))^2)
sst <- sum((grocery$y - mean(grocery$y))^2)
```

instead of dividing with the sigma we divide with sqrt(sum of square) to get 
unit length scaling

```{r}
grocery$x1 <- (grocery$x1 - mean(grocery$x1)) / sqrt(s11) 
grocery$x2 <- (grocery$x2 - mean(grocery$x2)) / sqrt(s22) 
grocery$x3 <- (grocery$x3 - mean(grocery$x3)) / sqrt(s33) 
grocery$y <- (grocery$y - mean(grocery$y)) / sqrt(sst)
```

as we are interested in independent variables get x columns
```{r}
w <- as.matrix(grocery[,c(1,2,3)]) 
```

calculate the covariance matrix W'W
```{r}
t(w) %*% w 
```

get the inverse of covariance matrix to get vif matrix where m_ii is the ith vif value
```{r}
solve(t(w) %*% w) #look at the diagonal values!
```

vifs are same! we have done it!! :)